---
title: "Student Mental Health Data Visualization"
author: "Melvis Onoriode"
date: '2022-05-19'
output:
  html_document: default
  pdf_document: default
  word_document: default
---  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=F}
library(tidyverse)
library(plotly)
library(readr)
library(ggplot2)
```

```{r}
health <- read.csv("Student Mental health.csv")
```

## Data Cleaning
```{r}
names(health) <- c('Timestamp', 'Gender', 'Age', 'Course', 'Year', 'CGPA', 'Married', 'Depression', 'Anxiety', 'Panic_Attack', 'Treatment')
indx <- apply(health, 2, function(x) any(is.na(x)))
indx
# There is missing data in our AGE column
which(is.na(health$Age))
# observation 44 -> since it is only one individual with missing data, we'd look to replace the missing with the median age of this group to keep the numbers clean rather than using a decimal as a mean.
health[44,'Age'] <- median(health$Age, na.rm =T)
```

# Data Visualization (Overview)



# Gender Distribution
```{r}
Health_SummaryStat <- health %>%
  group_by(Gender) %>%
  summarise(count = n(),
            percentage = round((n()/ nrow(health)), digits = 4))
Health_SummaryStat
colors <- c('rgb(211,94,96)','rgb(114,147,203)')
Gender_PieChart <- plot_ly(data = Health_SummaryStat, labels = ~Gender, values = ~percentage,
                type = 'pie', sort = F,
                textposition = 'inside',
                textinfo = 'label+percent',
                insidetextfont = list(color = 'White'),
                hoverinfo = 'text',
                text = ~count,
                marker = list(colors = colors,
                line = list(color = 'Black', width = 1)),
                showlegend = TRUE) 
Gender_PieChart <- Gender_PieChart %>% layout(title = 'Pie Chart of Gender')
Gender_PieChart
```
74.3% of observations were female compared to 25.7% 

# Depression 
```{r}
Health_SummaryStat2 <- health %>%
  group_by(Depression) %>%
  summarise(count = n(),
            percentage = round((n()/ nrow(health)), digits = 4))
Depression_PieChart <- plot_ly(data = Health_SummaryStat2, labels = ~Depression, values = ~percentage,
                type = 'pie', sort = F,
                textposition = 'inside',
                textinfo = 'label+percent',
                insidetextfont = list(color = 'White'),
                hoverinfo = 'text',
                text = ~count,
                marker = list(colors = colors,
                line = list(color = 'Black', width = 1)),
                showlegend = TRUE) 
Depression_PieChart %>% layout(title = 'Pie Chart of Depression')
```
34.6 % of sample had depression compared to an average 5.0% rate amongst adults in the population.

# Depression vs Gender

```{r}
health %>% 
  count(Gender, Depression, sort = F) %>%
  group_by(Gender) %>%
  mutate(prop = round((n / sum(n)),digits = 4)) %>%
  plot_ly(x = ~Gender, y=~prop, color = ~Depression, type = "bar",
          text = ~paste(Gender, prop*100 ,'%'),
          textposition = 'outside') %>%
    layout(barmode = 'Stacked',
           title = 'Barplot of Depression amongst Genders')
```
Females tended to on average have more cases of depression compared to male. (38.67% vs 23.08%)

# GPA Distribution
### Looking to see if there is a positive association between GPA and Depression (i.e higher CGPA -> see higher proportion depressed? Just initial thoughts)
```{r}
health$CGPA <- as.factor(health$CGPA)
levels(health$CGPA)
# Levels are in order ~
health %>%
  group_by(CGPA)%>%
  summarize(count = n()) %>%
  plot_ly(x =~CGPA, y=~count, type = 'bar',
        text = ~count,
        textposition = 'outside',
        marker = list(color = 'rgb(158,202,225)',
          line = list(color = 'black',
                 width = 1.0))) %>%
  layout(title = 'Distibution of CGPA')
```

# CGPA vs Depression
```{r}
health %>%
  count(CGPA, Depression, sort = F) %>%
  mutate(proportion = round((n/sum(n)),digits=4)) %>%
  plot_ly(x =~CGPA, y=~proportion, color = ~Depression, type = 'bar') %>%
  layout(barmode = 'Group',
         title = 'Barplot of Depression vs CGPA')
```

We aren't able to see any clear patterns with CGPA and Depression, but we do notice a non-monotonic increase in depression rate as CGPA increases. This might be due to inadequate group sizes, and if given a more percise CGPA value, we would of been able to get a clearer picture. But with this data set we do notice that CGPA has a slight positive associate with depression rate.

# Major Courses

### Since there are many different coures we'd look at the top ones and see if the depression rates vary amongst them.

```{r}
health %>% 
  group_by(Course) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  filter(count >2)
## Lets look at coures BCS, Engineering, BIT, Biomedical science , and KOE
```

# Top Major Courses vs Depression
```{r}
health %>%
  filter(grepl('BIT|KOE|BCS|Engineering|Biomedical science', Course)) %>%
  count(Course, Depression, sort = T) %>%
  group_by(Course) %>%
  mutate(prop = round((n / sum(n)),digits = 4)) %>%
  plot_ly(x = ~Course, y=~n, color = ~Depression, type = "bar",
          text = ~paste(Course, n),
          textposition = 'outside') %>%
  layout(barmode = 'Stacked',
         title = 'Barplot of Depression amongst the top 5 Courses')

```
Engineering students seem to have the most cases of depression, and surprisingly 0 came out of the Bio medical group. Group sizes too small to make any conclusive associations.
